Automatic Identification and Annotation of Tasks in Structured Programs

### 2018-02-02 docs/ACM/paper.tex
Talking a bit about task expansion.

\paragraph{Tasks: syntax and scope:}
Our goal is to identify {\em tasks} in {\em structured} programs.
Definition~\ref{def:task} formalizes the notion of task.
A structured program is a program that can be partitioned in {\em hammock
regions}, a concept introduced by Ferrante {\em et al.}~\cite{Ferrante87} in
the middle 80's.
The structure of this program is described by a a Control Flow Graph $G$ --
a directed graph with a starting node $s$, and an exit node $x$.
A hammock region is a subgraph $G' \subseteq G$, with a special node
$h$, which {\em dominates}\footnote{A node $n_1 \in G$ dominates another node
$n_2 \in G$ if every path from  $S$ to $n_2$ goes across $n_1$.
Inversely, $n_1$ post-dominates $n_2$ if every path from $n_2$ to $x$ must
go across $n_1$.} every other node in $G'$.
Additionally, there exists a node $w \in G$, $w \notin G'$, such that
$w$ {\em post-dominates} every node in $G'$.
In this definition, $h$ is the {\em entry point}, and $w$ is the {\em exit point}
of $G'$.

@@

% Explain the tension between large and small tasks.
The benefit of creating tasks depend on two factors:
runtime cost, and parallelism.
These two factors lead to a sweet spot in the size of tasks, where size is
measured as the amount of code they contain.
The smaller the tasks are, the more parallel they tend to be, as a consequence
of less dependencies.
However, if tasks are too small, then the performance gained by the extra
parallelism might not compensate the cost of managing them.
On the other hand, if tasks are too big, then we might not obtain enough
parallelism, because individual tasks still execute sequentially.
In this paper, we extract the smallest thread above the runtime cost.
To bestow meaning into this last statement, in what follows, we define more
formally the notions of task size and runtime cost.

@@

\begin{example}
\label{ex:cond_task}
The condition in line 7 of Figure~\ref{fig:ex_Regions} determines that tasks will
only be created if the expression \textsf{5 $\times$ M $<$ WORK\_CUTOOF}
holds.
The expression \textsf{5 $\times$ M} is an approximation for the size of the
task that comprises the loop at lines 8-10.
\end{example}

@@

% Discuss the expansion algorithm.
\paragraph{The Task Expansion Algorithm.}

>>>>

\paragraph{Tasks: syntax and scope:}
Our goal is to identify {\em tasks} in {\em structured} programs.
Definition~\ref{def:task} formalizes the notion of task.
A structured program is a program that can be partitioned in {\em hammock
regions}, a concept introduced by Ferrante {\em et al.}~\cite{Ferrante87} in
the middle 80's.
For completeness, we re-state it in Definition~\ref{def:hammock}.
The structure of this program is described by a a Control Flow Graph $G$ --
a directed graph with a starting node $s$, and an exit node $x$.

\begin{definition}[Hammock Region~\cite{Ferrante87}]
\label{def:hammock}
A hammock region $G'$ from a CFG $G$ is a subgraph $G' \subseteq G$, with a
node $h$, that {\em dominates}\footnote{A node $n_1 \in G$ dominates another node
$n_2 \in G$ if every path from  $S$ to $n_2$ goes across $n_1$.
Inversely, $n_1$ post-dominates $n_2$ if every path from $n_2$ to $x$ must
go across $n_1$.} the other nodes in $G'$.
Additionally, there exists a node $w \in G$, $w \notin G'$, such that
$w$ {\em post-dominates} every node in $G'$.
In this definition, $h$ is the {\em entry point}, and $w$ is the {\em exit point}
of $G'$.
\end{definition}

@@

% Explain the tension between large and small tasks.
The benefit of creating tasks depend on two factors:
runtime cost, and parallelism.
These two factors lead to a sweet spot in the size of tasks, where size is
measured as the amount of code they contain.
The smaller the tasks are, the more parallel they tend to be, as a consequence
of less dependencies.
However, if tasks are too small, then the performance gained by the extra
parallelism might not compensate the cost of managing them.
On the other hand, if tasks are too big, then we might not obtain enough
parallelism, because individual tasks still execute sequentially.
In this paper, {\em we extract the smallest thread above the runtime cost}.
To bestow meaning into this last statement, in what follows, we define more
formally the notions of task size and task expansion.

@@

\begin{example}
\label{ex:cond_task}
The condition in line 7 of Figure~\ref{fig:ex_Regions} determines that tasks will
only be created if the expression \textsf{5 $\times$ M $<$ WORK\_CUTOFF}
holds.
The expression \textsf{5 $\times$ M} is an approximation for the size of the
task that comprises the loop at lines 8-10.
The constant \textsf{WORK\_CUTOFF} represents the cost of creating a thread
in the OpenMP runtime, and it is determined empirically.
\end{example}

@@

% Discuss the expansion algorithm.
\paragraph{The Task Expansion Algorithm.}
Task expansion consists in finding the smallest task that is large enough to
pay for the thread creation cost.
Figure \ref{fig:expand_alg} shows the algorithm that performs this activity.
We start this algorithm by setting \textsf{REGION} to be a hammock graph $H$
that is contained entirely within a windmill $W$.
The hammock decomposition of a structured program forms a tree~\cite{Ferrante87},
such that vertices in this tree represent hammock graphs.
$H_1$ is a child of $H_2$ in this tree if $H_1 \subset H_2$, and for any other
node $H_x$, $H_1 \subset H_x \Rightarrow H_2 \subset H_x$.
In this context, we call $H_2$ the \textsf{PARENT} of $H_1$.
